---
title: "Homework 2"
author: "Jeffrey Liang"
date: "9/23/2020"
output: 
 github_document
---
```{r setup ,echo = F, message= F}
library(tidyverse)
```

# Problem 1 

## Load Wheel data
```{r load_whell_data, warning=F, message=FALSE}
data_collection = c() #Load the wheel data
for (i in c(1:3)){
  data_collection[[i]] = 
    readxl::read_excel(
      paste(here::here(),"/data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sep=''),
      skip = 1, 
      sheet = i
    ) %>% 
    janitor::clean_names() %>% 
    select( -starts_with('X1'))
}

wheel_df = bind_rows(data_collection)

preci_collection = c() # load the precipitation data in wheel
for (i in c(4:9)){
  preci_collection[[i]] = 
    readxl::read_excel(
      paste(here::here(),"/data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sep=''),
      skip = 1,
      sheet = i
    ) %>% 
    janitor::clean_names() %>% 
    mutate(year = 2020 + 3 -i)
}

preci_df = bind_rows(preci_collection)

rm(data_collection,preci_collection)
```


## Clean Wheel data

```{r clean_data}
wheel_df = #clean wheel
  wheel_df %>% 
  filter(
    #!grep(" Total",get('month')), #this command return vector with null row, so the data can't subset
    !str_detect(get("month")," Total"),
    !is.na(get("dumpster")) #Or drop_na(dumpster)
    ) %>% 
  mutate(sports_balls = 
           as.integer(round(get("sports_balls")))
         )
tail(wheel_df)
```

## Clean precipitation data
```{r}
preci_df = 
  preci_df %>%
  filter(!is.na(get("month")),
         get('year') %in% c(2018,2017)
         ) %>% 
  mutate(month =
           apply( # this is the apply() function in R, need to specific dim before giving function, weird
             as.matrix(           # as.matrix is very important here, cuz pull() or get() have NULL dim()
             get("month")
             ), 1, # specific dim()
             function(x) month.name[[x]]) # this is how to write lambda in R, clean
  ) %>%  # OR use month = month.names[month], or create a tabel with order and then left join
  relocate(year)
tail(preci_df)
```


## Write something about this data
 
\newpage
# Problem 2

## load data
```{r load_nyc,message=F,warning=F}
rm(list=ls())
read_nyc = function(){
return_data = read_csv(
  paste(here::here(),"/data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",sep = '')
) %>% 
  janitor::clean_names()
return(return_data)
}
#skimr::skim_without_charts(nyc_df) # TODO: route* need to be in the same columns
```

## Clean data
```{r how_to_do_for_loop_mutation, include=FALSE,echo = FALSE}
nyc_df = read_nyc()
for (i in 1:11){
  col_name = paste("route",i,sep='')
  mutate(nyc_df,!!as.symbol(col_name) := as.character(pull(nyc_df,as.name(col_name))))
  #use !! before variable to avoid evaluation, use := for dynamic name assigning
  # I have to say, dplyr is readable, but pain in the ass from a computing perspective,
  # one thing for sure that it would be great amount of work transferring dplyr script to other language
}
```

```{r clean_nyc_data,warning=F,message=F}
nyc_df = 
  read_nyc() %>% 
  select(line:station_longitude,route1:route11,vending,entry,entrance_type,ada) %>% 
  mutate(entry = if_else(
    apply(as.matrix(entry),1,str_to_lower)=="yes",
    TRUE,FALSE,NA),
    vending = if_else(vending=="YES",T,F,NA),
    #across(c("line","station_name"),function(x) str_replace_all(str_to_lower(x)," ",""))
    )

nyc_df_tidy = nyc_df %>% 
  mutate(across(matches('route'),as.character)
         )%>% #OR use across(matches("something",fun))
  relocate(route1:route11,.after = last_col()) %>% 
  pivot_longer( # clean route* variables to meaningful route variable
    cols = route1:route11, #route 8 is numeric
    names_to = "route_name",
    values_to = 'route_number',
    names_prefix = 'route'
  ) %>% 
  drop_na(route_number)
tail(nyc_df)
```
This NYC's metro system data is a `r nrow(read_nyc())` x `r ncol(read_nyc())` size data, describing NYC metro system's **_`r paste(names(read_nyc()),sep = ',')`_**.

The data is not beautiful in all sense: 

1. It has route all over the place and name and line separated instead of treating as unique indentity;
1. value types in route* were not consistent;
1. Boolean variable in _vending_ and _entry_ were in form of "YES"/"NO" instead of boolean.

So the first step I did to the data was to select required variables. Followed with changing values in _entrance_ variable to boolean factors with if_else() function. Then I took the information in the route\*: first dealed with route 8 to 11 which not consistent with the other line format of data and put them into *route_name* and *route_number* variables using pivot_longer(). To seperate 

The outcome data after piping is a dataset of `r nrow(nyc_df)` x `r ncol(nyc_df)` size dataset, with variables of **_`r names(nyc_df)`_**, and the data with following properties:

* there's `r count(distinct(nyc_df,station_name,line))` unique station (including names and line) in this data;
* Of `r nrow(nyc_df)` total stations, `r sum(pull(nyc_df,'ada'))` are ADA compliant stations;
* Of `r nrow(filter(nyc_df,!vending))` stations without vending, 
`r round(nrow(filter(nyc_df,!vending,entry))/nrow(filter(nyc_df,!vending)),2)` stations' entrance/exits without vending allow entrance.

After transforming the NYC transit data, of `r count(distinct(nyc_df_tidy,station_name,line))` stations, 
`r nyc_df_tidy %>% distinct(station_name,line,.keep_all=TRUE) %>% filter(route_number=="A") %>% count()` serve route A. Among these stations, `r nyc_df_tidy %>% distinct(station_name,line,.keep_all=TRUE) %>% filter(route_number=="A",ada) %>% count()` are ADA compliant.

\newpage
# Problem 3
## Load data
### Pols Month Data

```{r load_five, message=F,warning=F}
rm(list=ls())
read_five = function(data_name_index = 1){
  data_name = str_c(c("pols-month",'unemployment',"snp"),".csv")
  read_data = 
    read_csv(paste(here::here(),"/data/fivethirtyeight_datasets/",data_name[[data_name_index]],sep=''),
             ) %>% 
    janitor::clean_names()
}

#load month data
pols_month = read_five(1) %>% 
  separate(mon,c("year","month",'day'),"-") %>% 
  mutate(across(.cols = year:day,as.integer)) %>% 
  left_join(tibble(month = 1:12,month_name = month.name)) %>% 
  mutate(month = month_name) %>% 
  select(-month_name) %>% 
  pivot_longer( # OR mutate president = 
    c("prez_gop","prez_dem"),
    names_to = "president",
    values_to = "president_boolean",
    names_prefix="prez_"
  ) %>% 
  filter(president_boolean == 1) %>% 
  select(-c(president_boolean,day)) %>% 
  mutate(president = factor(president,levels = c('dem',"gop")))
str(pols_month)
```

First I loaded the Pols Month Data, the original data countains *`r read_five(1) %>% names() %>% paste(sep=',')`* columns, in which *prez_dem, prez_gov* represent an indicator columns of
which parties' candidates were in charge of White House. The columns *gov_dem,sen_dem,rep_dem*
 represent the number of governors, senators and representatives of Democrate Party and the others represent those of GOP. The data was record in the date of count on *date*. The data is in `r paste(nrow(read_five(1))," x ",ncol(read_five(1)))` dimension.

\ To make this data looks/works more tidy, I did the following: first I seperated *date* into _year,month,day_ ,changing the _month_ value into full name instead of number and keep the first two columns only. Second, *prez_dem, prez_gop* columns were melted together, keeping the *dem, gop* information in column names, their values to an indicator columns. 
And use filter to the indicator columns to selected the informative rows( that is president's party of the record date). And the final data is in `r paste(nrow(pols_month)," x ",ncol(pols_month))` dimension.


### SNP dataset
```{r load_and_clean_snp_data, message=F,warning=F}
snp_data = read_five(3) %>% 
  separate(col = date,into = c("month","day","year"),"/") %>% 
  select(-day) %>% 
  mutate(across(.cols = year:month,as.numeric),
         month = month.name[month]
         ) %>% 
  relocate(year,month)
str(snp_data)
```

Then the SNP dataset is loaded, same workflow was process: separating _date_ in to _month,day,year_, translate the numeric _month_ variable to character full name and drop _day_ as required. After workflow, a `r nrow(snp_data)` x `r ncol(snp_data)` dimension data is produced.

### Unemployment dataset
```{r load_and_clean_unemployment_data, message=F,warning=F}
unemploy_data = read_five(2) %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to="unemploy_rate"
  ) %>% 
  mutate(month = apply(as.matrix(month),1,
                       function(x) month.name[[
                         match(x,str_to_lower(month.abb))
                         ]]))
str(unemploy_data)
```

Finally the Unemployment dataset was taken and similar workflow implemented. First columns of _`r select(read_five(2),jan:dec) %>% names()`_, containing information of unemployment rate in the month, were melted into variable _month_ and  _unemploy_rate_. Then the abbreviated month name was translated into full name for consistence. An `r nrow(unemploy_data)` * `r ncol(unemploy_data)` is produced.

## Join data
```{r}
five30eight = pols_month %>% 
  left_join(snp_data, by = c("year","month")) %>% 
  left_join(unemploy_data, by = c("year","month"))
skimr::skim_without_charts(five30eight)
```

